---
title: "Buidling Geom Hurricane"
author: "vh42720"
date: "November 25, 2017"
output:  md_document
---

## Introduction

The following is the quick walk-through for building a hurricane geom as the final assignment of John Hopskin's Building Data Visualizaton Tools course on Coursera. My purpose is to provide a quick guide to understand the problem that many are struggling with.

Loading the following packages for the project

```{r loading packages, warning=FALSE, message=FALSE}
library(ggplot2)
library(ggmap)
library(tidyverse)
library(lubridate)
library(geosphere)
library(grid)
source("build_geom.R")
ike2 <- readRDS("ike2")
```

## The task

The purpose of this assignment is to draw on your knowledge of the grid and ggplot2 package to build a new geom. You will then need to load and tidy the provided dataset in order to plot the geom on a map. In short you will build this:

```{r example, warning=FALSE, message=FALSE}
map_plot <- ggmap::get_map("Lousiana", zoom = 5, maptype = "toner-background") 

map_plot %>%
    ggmap::ggmap(extent = "device") +
    geom_hurricane(data = ike2,
                   ggplot2::aes(x = longitude, y = latitude, 
                                r_ne = ne, r_se = se, r_nw = nw, r_sw = sw,
                                color = wind_speed, fill = wind_speed)) +
    ggplot2::scale_color_manual(name = "Wind speed (kts)", 
                                values = c("red", "orange", "yellow")) + 
    ggplot2::scale_fill_manual(name = "Wind speed (kts)", 
                               values = c("red", "orange", "yellow"))
```

## Breaking down the problems

While the problem seems complicated at first glance (and it is), breaking it down to few simple steps can help you understand it right away.

1. Tidy the data to have the same columns as shown. Change the data type as you see fit. Notice the order of the last 4 columns: it is not in the order of ne -> se -> sw -> nw yet. BUT, you have to build the geom in the right order by passing it to the required_aes as the right order or else it will be a mess to fix later on!

```{r tidy data}
ike2
```

2. The longitude (negative) and latitude in the above table will be the center of the graph. The values in ne, se, sw, nw columns are radius of respective the quadrants after they are transform into the nautical miles to use on map. This is nothing more than multiply the value by 1852.

```{r nautical miles conversion}
data.frame(old = ike2$ne, new = ike2$ne * 1852)
```

3. To create the graph, it is nothing more than connecting the points. Specifically, you use the center (lat & long) and radius (ne, se, sw, nw) to find all the points (in longitude and latitude) to create each quadrant circle. geosphere::destPoint will make this easy. NOTICE the order of the direction. We are going around a circle from left to right!!!

As an example, lets say I take the first row of the above table and plot the points for the first circle.

```{r geosphere destpoint}
temp <- ike2[1,]
arc <- data.frame(start_arc = c(0, 90, 180, 270),
                  end_arc = c(90, 180, 270, 360))
# Transform to long data
temp <- temp %>%
    gather("direction", "wind_radius", c("ne", "se", "sw", "nw"))%>%
    bind_cols(arc)

## Create the data frame of all points
points <- data.frame()
for(i in 1:nrow(temp)){
    out <- geosphere::destPoint(p = c(temp[i,]$longitude, temp[i,]$latitude),
                         b = temp[i,]$start_arc:temp[i,]$end_arc,
                         d = temp[i,]$wind_radius)
    out <- as.data.frame(out)
    names(out) <- c("longitude", "latitude")
    points <- bind_rows(points, out)
}
head(points)

plot(points)
```

Now you see destPoint in action, the rest is to connect them with polygonGrob and you are done. 

4. The last point is how to put all these ideas into the geom. Remember to create a geom you create 2 things: a geom class (lets call it GeomHurricane) and a wrapper geom call (geom_hurricane). The geom class will handle everything and the geom call you just need to copy exactly like in the book with geom = GeomHurricane. 

#### Important

There are 6 required aes (x, y, r_ne, r_se, r_sw, r_nw) passed to the call (and thus the class). This is nothing more than creating a dataframe of all these required columns which are all you need to build the graph. 

Secondly, draw_group will AUTOMATICALLY loop over each row of the data so you will input the code to work with ONE ROW only. (Read the build_geom.R for more)

## Final Notes

Trying out destPoint for yourself will help you tremendously to understand the problem. Moreover, don't get angry if you are stuck on something. Relax and start reading the documentation of what you don't understand. It is a very good learning experiment for my part. 

And lastly, have fun!